Mathematical description of Naomi

### 2022-12-07 src/example_naomi/maths.Rmd
A little more while travelling

* Let $\nu$ be a household survey occuring at $T_1$
* Furnishes weighted observations for HIV prevalence $\hat \rho_{\{x\}, \{s\}, \{a\}, \nu}$, ART coverage $\hat \alpha_{\{x\}, \{s\}, \{a\}, \nu}$ and proportion recently infected $\hat \kappa_{\{x\}, \{s\}, \{a\}, \nu}$ with respective Kish effective sample sizes $M^{\hat \theta}_{\{x\}, \{s\}, \{a\}, \nu}$ for $\theta \in \{\rho, \alpha, \kappa\}$
* The observed cases are $Y^{\hat \theta}_{\{x\}, \{s\}, \{a\}, \nu} = M^{\hat \theta}_{\{x\}, \{s\}, \{a\}, \nu} \cdot \hat \theta_{\{x\}, \{s\}, \{a\}, \nu}$
* For HIV prevalence we use a binomial working likelihood
\begin{align}
Y^{\hat \rho}_{\{x\}, \{s\}, \{a\}, \nu} &\sim \text{Bin}(M^{\hat \rho}_{\{x\}, \{s\}, \{a\}, \nu}, \rho_{\{x\}, \{s\}, \{a\}, {T_1}}) \\
\rho_{\{x\}, \{s\}, \{a\}, t} &= \frac{\sum_{x \in \{x\}} \sum_{s \in \{s\}} \sum_{a \in \{a\}} N_{x, s, a, t} \cdot \rho_{x, s, a, t}}{\sum_{x \in \{x\}} \sum_{s \in \{s\}} \sum_{a \in \{a\}} N_{x, s, a, t}}
\end{align}
* For ART coverage we also use a binomial working likelihood
\begin{align}
Y^{\hat \alpha}_{\{x\}, \{s\}, \{a\}, \nu} &\sim \text{Bin}(M^{\hat \alpha}_{\{x\}, \{s\}, \{a\}, \nu}, \alpha_{\{x\}, \{s\}, \{a\}, {T_1}}) \\
\alpha_{\{x\}, \{s\}, \{a\}, t} &= \frac{\sum_{x \in \{x\}} \sum_{s \in \{s\}} \sum_{a \in \{a\}} N_{x, s, a, t} \cdot \rho_{x, s, a, t} \cdot \alpha_{x, s, a, t}}{\sum_{x \in \{x\}} \sum_{s \in \{s\}} \sum_{a \in \{a\}} N_{x, s, a, t} \cdot \rho_{x, s, a, t}}
\end{align}
* For recent infections we also use a binomial working likelihood
\begin{align}
Y^{\hat \kappa}_{\{x\}, \{s\}, \{a\}, \nu} &\sim \text{Bin}(M^{\hat \kappa}_{\{x\}, \{s\}, \{a\}, \nu}, \kappa_{\{x\}, \{s\}, \{a\}, {T_1}}) \\
\kappa_{x, s, a, t} &= 1 - \exp(- \lambda_{x, s, a, t} \cdot \frac{1 - \rho_{x, s, a, t}}{\rho_{x, s, a, t}} \cdot (\Omega_T - \beta_T ) - \beta_T ) \\
\Omega_T &\sim \mathcal{N}(\Omega_{T_0}, \sigma^{\Omega_T}) \\
\beta_T &\sim \mathcal{N}(\beta_{T_0}, \sigma^{\beta_T})
\end{align}

>>>>

* Let $\nu$ be a household survey occuring at $T_1$
* Furnishes weighted observations for HIV prevalence $\hat \rho_{\{x\}, \{s\}, \{a\}, \nu}$, ART coverage $\hat \alpha_{\{x\}, \{s\}, \{a\}, \nu}$ and proportion recently infected $\hat \kappa_{\{x\}, \{s\}, \{a\}, \nu}$ with respective Kish effective sample sizes $M^{\hat \theta}_{\{x\}, \{s\}, \{a\}, \nu}$ for $\theta \in \{\rho, \alpha, \kappa\}$
* The observed cases are $Y^{\hat \theta}_{\{x\}, \{s\}, \{a\}, \nu} = M^{\hat \theta}_{\{x\}, \{s\}, \{a\}, \nu} \cdot \hat \theta_{\{x\}, \{s\}, \{a\}, \nu}$
* For HIV prevalence we use a binomial working likelihood
\begin{align}
Y^{\hat \rho}_{\{x\}, \{s\}, \{a\}, \nu} &\sim \text{Bin}(M^{\hat \rho}_{\{x\}, \{s\}, \{a\}, \nu}, \rho_{\{x\}, \{s\}, \{a\}, {T_1}}) \\
\rho_{\{x\}, \{s\}, \{a\}, t} &= \frac{\sum_{x \in \{x\}} \sum_{s \in \{s\}} \sum_{a \in \{a\}} N_{x, s, a, t} \cdot \rho_{x, s, a, t}}{\sum_{x \in \{x\}} \sum_{s \in \{s\}} \sum_{a \in \{a\}} N_{x, s, a, t}}
\end{align}
* For ART coverage we also use a binomial working likelihood
\begin{align}
Y^{\hat \alpha}_{\{x\}, \{s\}, \{a\}, \nu} &\sim \text{Bin}(M^{\hat \alpha}_{\{x\}, \{s\}, \{a\}, \nu}, \alpha_{\{x\}, \{s\}, \{a\}, {T_1}}) \\
\alpha_{\{x\}, \{s\}, \{a\}, t} &= \frac{\sum_{x \in \{x\}} \sum_{s \in \{s\}} \sum_{a \in \{a\}} N_{x, s, a, t} \cdot \rho_{x, s, a, t} \cdot \alpha_{x, s, a, t}}{\sum_{x \in \{x\}} \sum_{s \in \{s\}} \sum_{a \in \{a\}} N_{x, s, a, t} \cdot \rho_{x, s, a, t}}
\end{align}
* For recent infections we also use a binomial working likelihood
\begin{align}
Y^{\hat \kappa}_{\{x\}, \{s\}, \{a\}, \nu} &\sim \text{Bin}(M^{\hat \kappa}_{\{x\}, \{s\}, \{a\}, \nu}, \kappa_{\{x\}, \{s\}, \{a\}, {T_1}}) \\
\kappa_{x, s, a, t} &= 1 - \exp(- \lambda_{x, s, a, t} \cdot \frac{1 - \rho_{x, s, a, t}}{\rho_{x, s, a, t}} \cdot (\Omega_T - \beta_T ) - \beta_T ) \\
\Omega_T &\sim \mathcal{N}(\Omega_{T_0}, \sigma^{\Omega_T}) \\
\beta_T &\sim \mathcal{N}(\beta_{T_0}, \sigma^{\beta_T})
\end{align}
  * $\kappa_{x, s, a, t}$ are the predicted proportion recently infected among HIV positive persons
  * $\Omega_T$ is the mean duration of recent infection (MDRI)
    * Use an informative prior based on the characteristics of the recent infection testing algorithm (RITA)
    * For PHIA surveys $\Omega_{T_0} = 130 \text{ days}$ and $\sigma^{\Omega_T} = 6.12 \text{ days}$
    * **Flag: what about for other surveys?*
  * $\beta_T$ is the false recent ratio (FRR)
    * For PHIA surveys $\beta_{T_0} = 0.0$ and $\sigma^{\beta_T} = 0.0$. **Flag: is this just assuming that there is no false recency? How is this assumption justified?*

## ANC testing data

* Include two years of ANC testing data: the year of the most recent survey $Y[T_1]$ and the current year $Y[T_2]$
* $W^\text{ANC}_{\{x\}, Y[t]}$ are the number of ANC clients
* $X^\text{ANC}_{\{x\}, Y[t]}$ are the number of ANC clients with ascertained status
* $Y^\text{ANC}_{\{x\}, Y[t]}$ are the number of ANC clients with positive status (either known or tested)
* $Z^\text{ANC}_{\{x\}, Y[t]}$ are the number of ANC clients already on ART prior to first ANC
* Sometimes ANC testing data are only available for part of a given year, for example only the first three quarters. Denote $M^\text{ANC}_{Y[t]} \in \{1, \ldots, 12\}$ the number of months of reported data reflected in counts for year $Y[t]$
